---
title: "Validation Set Approach - HW3"
output: html_notebook
---

We will estimate the test error of logistic regression model using the validation set approach. 

```{r}
library(ISLR)
attach(Default) # makes Auto variables available by name

set.seed(1)
```

(a) Fit a logistic regression model that uses income and balance to predict default.

```{r}
glm.fit <- glm(default~income+balance, Default, family="binomial") # fit logistic regression
summary(glm.fit)

```


(b) Using the validation set approach, estimate the test error of this model. In order to do this, you must perform the following steps:

i.) Split the sample set into a training set and a validation set.

```{r}
train = sample(dim(Default)[1], size = 0.50*dim(Default)[1], replace=FALSE)	# set 50% of data set as training set
val_set = -train 				# set remaining 50% of data set as validation set
training_data = Default[train, ]		# split training set

val_data= Default[val_set, ]		# split test set

dim(Default) # entire data set
dim(training_data) # training set
dim(val_data) # Validation set
```

ii.) Fit a multiple logistic regression model using only the training observations.

```{r}
glm.fit <- glm(default~income+balance, Default, family="binomial", subset = train)
summary(glm.fit)
```

iii.) Obtain a prediction of default status for each individual in the validation set by computing the posterior probability of default for that individual, and classifying the individual to the default category if the posterior probability is greater than 0.5.

```{r}
glm.probs = predict(glm.fit, val_data, type="response") # compute posterior prob
glm.pred = ifelse(glm.probs>0.5, "Yes", "No") # classifying 
```

iv.) Compute the validation set error, which is the fraction of the observations in the validation set that are misclassified.

```{r}
table(glm.pred, val_data$default) #accuracy
mean(glm.pred != val_data$default) # validation error
```

(c) Repeat the process in (b) three times, using three different splits of the observations into a training set and a validation set. Comment on the results obtained.

```{r}
train = sample(dim(Default)[1], size = 0.50*dim(Default)[1], replace=FALSE)	# set 50% of data set as training set
val_set = -train 				# set remaining 50% of data set as validation set
training_data = Default[train, ]		# split training set
val_data= Default[val_set, ]		# split validation set
glm.fit <- glm(default~income+balance, Default, family="binomial", subset = train)
glm.probs = predict(glm.fit, val_data, type="response") # compute posterior prob
glm.pred = ifelse(glm.probs>0.5, "Yes", "No") #  classifying the individual to the default category if the posterior probability is greater than 0.5
table(glm.pred, val_data$default) 
mean(glm.pred != val_data$default) # validation error

train = sample(dim(Default)[1], size = 0.50*dim(Default)[1], replace=FALSE)	# set 50% of data set as training set
val_set = -train 				# set remaining 50% of data set as validation set
training_data = Default[train, ]		# split training set
val_data= Default[val_set, ]		# split validation set
glm.fit <- glm(default~income+balance, Default, family="binomial", subset = train)
glm.probs = predict(glm.fit, val_data, type="response") # compute posterior prob
glm.pred = ifelse(glm.probs>0.5, "Yes", "No") #  classifying the individual to the default category if the posterior probability is greater than 0.5
table(glm.pred, val_data$default) 
mean(glm.pred != val_data$default) # validation error


train = sample(dim(Default)[1], size = 0.50*dim(Default)[1], replace=FALSE)	# set 50% of data set as training set
val_set = -train 				# set remaining 50% of data set as validation set
training_data = Default[train, ]		# split training set
val_data= Default[val_set, ]		# split validation set
glm.fit <- glm(default~income+balance, Default, family="binomial", subset = train)
glm.probs = predict(glm.fit, val_data, type="response") # compute posterior prob
glm.pred = ifelse(glm.probs>0.5, "Yes", "No") #  classifying the individual to the default category if the posterior probability is greater than 0.5
table(glm.pred, val_data$default) 
mean(glm.pred != val_data$default) # validation error
```


(d) Now consider a logistic regression model that predicts the probability of default using income, balance, and a dummy variable for student. Estimate the test error for this model using the validation set approach. Comment on whether or not including a dummy variable for student leads to a reduction in the test error rate.

```{r}
dummy_student = ifelse(student=="Yes", 1, 0) # create dummy variable for student
head(Default)
Default = cbind(Default, dummy_student)
head(Default)

train = sample(dim(Default)[1], size = 0.50*dim(Default)[1], replace=FALSE)	# set 50% of data set as training set
val_set = -train 				# set remaining 50% of data set as validation set
training_data = Default[train, ]		# split training set
val_data= Default[val_set, ]		# split validation set

glm.fit <- glm(default~income+balance+dummy_student, Default, family="binomial", subset = train)
glm.probs = predict(glm.fit, val_data, type="response") # compute posterior prob
glm.pred = ifelse(glm.probs>0.5, "Yes", "No") #  classifying the individual to the default category if the posterior probability is greater than 0.5
table(glm.pred, val_data$default) 
mean(glm.pred != val_data$default) # validation error

```

8. We will now perform cross-validation on a simulated data set.

(a) Generate a simulated data set as follows:
```{r}
set.seed(1)
x=rnorm(100)
y=x-2*x^2+rnorm (100) 
```


```{r}
plot(x,y)
```

