---
title: "Linear Discriminant Analysis (LDA)"
output:
  html_document:
    df_print: paged
---

```{r}
require(ISLR) # load package
require(MASS) # load package
```

```{r}
head(Smarket) # stock market data set
```
Use stock market data to show the Direction the market took on a particular day. The response is going to be the direction the market took on a particular day. And we'll use the returns before 2005 (train set) to try and predict the direction in 2005 (test set).

```{r}
lda.fit = lda(Direction~Lag1+Lag2, Smarket, subset=Year<2005)
```


```{r}
lda.fit
```
The probabilities of groups show that the market is volatile. More precisely, about 50% of the time the marker is up (.51) and down(.49).

The command below plots the linear discriminant function separately for the Up group and the down group.

```{r}
plot(lda.fit)
```

Predict stock market in year 2005 (aka test set).
```{r}
Smarket.2005 = subset(Smarket, Year==2005) # Create a new data frame for year 2005
```
```{r}
Smarket.2005
```

```{r}
lda.pred = predict(lda.fit, Smarket.2005)
```

```{r}
data.frame(lda.pred)[1:5,]
```
```{r}
table(lda.pred$class, Smarket.2005$Direction)
```

```{r}
mean(lda.pred$class== Smarket.2005$Direction)
```
This means our prediction is 55% accurate